\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)

### 2015-10-19 leslieandkenny.Rnw
Tried some more glmms

\title{\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)}
\author{Leslie Gains-Germain, Kenny Flagg}
\date{Spring 2015}

@@

@@

# Combine nitrogen and inoc/cntl (for the heatmap labels)
#fert2$n.status <- with(fert2, interaction(n.trt, status))
@

@@

\begin{figure}[h!]
\caption{Example \texttt{itableplot} output.}
<<tabplot,echo=FALSE,warning=FALSE,message=FALSE>>=
tableplot(fert.sim1, select = c(infected, variety, block, n.trt, status),
          nBins = 30, sortCol = "variety")
@
\end{figure}

@@

>>>>

\title{\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)}
\author{Leslie Gains-Germain, Kenny Flagg}
\date{Fall 2015}

@@

# This forces the constraint that the block effects sum to 0
contrasts(fert2$block) <- contr.sum(6, contrast=TRUE)

@@

# Create a plot factor
fert2$plot <- factor(1:180)

# Combine nitrogen and inoc/cntl (for the heatmap labels)
fert2$n.status <- with(fert2, interaction(n.trt, status))
@

@@

\begin{figure}[h!]
\caption{Example \texttt{itableplot} output.}
<<tabplot,echo=FALSE,warning=FALSE,message=FALSE>>=
#Warning message:
#In tableplot_checkCols(substitute(sortCol), colNames) :
#  Sorting on multiple columns not supported anymore. Only first column is taken.
tableplot(fert.sim1, select = c(infected, variety, block, n.trt, status),
          nBins = 30, sortCol = "variety")
@
\end{figure}

@@

<<glmms,echo=FALSE,warning=FALSE,message=FALSE,results='hide',cache=TRUE>>=
glmBin <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
              family = binomial, data = fert2)
# Warning message:
# glm.fit: fitted probabilities numerically 0 or 1 occurred
# Residual deviance: 427.939  on 125  degrees of freedom
# 427.939 / 125 = 3.423512

glmQBin <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
               family = quasibinomial, data = fert2)
# No warnings
# Residual deviance: 427.939  on 125  degrees of freedom
# (Dispersion parameter for quasibinomial family taken to be 3.2817219)

glmmVar <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                   (1|block/variety),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# variety:block (Intercept) 0.46633  0.68289
# block         (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.53042 2535.76041
# varietySNMS:status2CNTL                    -18.50610 2527.55827
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11619 2527.55834
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.07429 4381.94303
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmVarPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                       (1|block/variety) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In vcov.merMod(object, use.hessian = use.hessian) :
#   variance-covariance matrix computed from finite-difference Hessian is
# not positive definite or contains NA values: falling back to var-cov estimated from RX
# 2: In vcov.merMod(object, correlation = correlation, sigm = sig) :
#   variance-covariance matrix computed from finite-difference Hessian is
# not positive definite or contains NA values: falling back to var-cov estimated from RX
# plot          (Intercept) 1.03818  1.0189
# variety:block (Intercept) 0.28324  0.5322
# block         (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.10372  8008.61084
# varietySNMS:status2CNTL                     -19.96154  7604.01229
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.93076  7604.01238
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.61822 13745.14025
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmRow <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                   (1|block/row),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# row:block (Intercept) 0.46633  0.68289
# block     (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.52017 2522.79684
# varietySNMS:status2CNTL                    -18.50714 2528.86662
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11723 2528.86668
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.06600 4374.54075
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmRowPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                       (1|block/row) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# plot      (Intercept) 1.03818  1.0189
# row:block (Intercept) 0.28324  0.5322
# block     (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.06813  7867.38080
# varietySNMS:status2CNTL                     -19.95030  7561.42603
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.91952  7561.42612
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.67575 13394.17908
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|plot),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning message:
# In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model is nearly unidentifiable: large eigenvalue ratio
#  - Rescale variables?
# plot   (Intercept) 1.3056   1.1426
# varietySNMS:n.trtLATE SPRING              -19.20256  371.85490
# varietySNMS:status2CNTL                   -19.21857  211.30941
# varietySNMS:n.trtEARLY SPRING:status2CNTL  20.44165  211.30904
# varietySNMS:n.trtLATE SPRING:status2CNTL   19.91369  260.18702
# deviance df.resid
#    731.7      149
# 731.7 / 149 = 4.9107383

glmmVarS <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                    (variety|block),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# block  (Intercept) 0.62664  0.79160
#        varietyPRHN 0.98385  0.99189  -0.933
#        varietySNMS 3.09082  1.75807  -0.713  0.876
#        varietyTAM  0.38672  0.62186  -0.471  0.432  0.162
#        varietyYSTN 0.86358  0.92929  -0.798  0.687  0.277  0.806
# varietySNMS:n.trtLATE SPRING               -18.54752 2546.52186
# varietySNMS:status2CNTL                    -18.51916 2537.53981
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.13402 2537.53988
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.08471 4406.80309
# deviance df.resid
#    842.8      135
# 842.8 / 135 = 6.2422222

# Compare coefs and SEs
cbind(coef(glmQBin)[c(1,7:13,34:55)], fixef(glmmVar))
cbind(arm::se.coef(glmQBin)[c(1,7:13,34:55)], arm::se.fixef(glmmVar))
# Most of the quasibinomial SEs are less than twice as big as the glmm SEs

# Now with plot-level REs
cbind(coef(glmQBin)[c(1,7:13,34:55)], fixef(glmmVar))
cbind(arm::se.coef(glmQBin)[c(1,7:13,34:55)], arm::se.fixef(glmmVar))
# Pretty similar. Maybe this is ok?
@
